R Notebook

### 2022-03-04 Code/Simulation.Rmd
Added copula ks-test histogram, added babu to fitted parameters, removed block bs sections

```{r}
# Load packages
library(xtable)
library(dgof)
library(tidyverse)
library(KSgeneral)
library(snpar)
```

@@

  # Parametric bootstrap to approximate the null distribution
  n <- length(x)
  stat.b <- double(B)
  for (i in 1:B) {
    x.b <- rnorm(n, mu, sigma)
    mu.b <- mean(x.b)
    sigma.b <- sd(x.b)
    stat.b[i] <- ks.test(x.b, "pnorm", mu.b, sigma.b)$statistic
    # write own ks test statistic formula here
    # is.character(y)
  }
  p.value <- (sum(stat.b >= stat) + 0.5) / (B + 1)

  list(p.value = p.value)
}

@@

# Generate results
pvals_parametric <- replicate(100, test_parametric(rnorm(100), B = 1000)$p.value)
pvals_naive <- replicate(100, test_naive(rnorm(100))$p.value)

# Create a master data_frame with all the results
histogram <- c(rep("Parametric", 100), rep("Naive", 100))
p_values <- c(pvals_parametric, pvals_naive)
results <- data_frame(p_values, histogram)

# Generate histograms plot
hist_parametric <- ggplot(results) + 
                   geom_histogram(aes(x = p_values, fill = histogram), boundary = 0, binwidth = 1/20) + 
                   scale_x_continuous(name = "p-value", breaks=c(0, 1)) + 
                   theme(legend.position = "none") +
                   facet_wrap(~histogram, scales = "free_y")

# Save the histogram figure
ggsave(filename = "hist_parametric.pdf", plot = hist_parametric, path = "../Manuscript", height = 2.1, width = 6)
```

```{r block functions}
# Functions supporting block bootstrap

# Add small noise to data
add_noise <- function(x) {

  # Grab variance and length
  v <- var(x)
  n <- length(x)

  # Generate noise where standard deviation is 1% of variance
  noise <- rnorm(n, mean = 0, sd = 1/100*v)

  # Return jittered data
  return(x + noise)
}

# Non-moving block generation
nmblk <- function(n, l) {

  # Number of blocks
  b <- n / l

  # Create non-moving block indices
  idx <- sample(seq(1, n - l + 1, by = l), b, replace = TRUE)

  # Return indices
  c(outer(0:(l - 1), idx, "+"))
}

# Moving block generation
mvblk <- function(n, l) {

  # Number of blocks
  b <- n / l

  # Create moving block indices
  idx <- sample(1:(n - l + 1), b, replace = TRUE)

  # Return indices
  c(outer(0:(l - 1), idx, "+"))
}
```

```{r hist_block}
# Block bootstrap function
blkboot <- function(tseries, statistic, R, l) {

  n <- length(tseries)

  # Gather observed statistic
  stat <- statistic(tseries)

  # Replicate block bootstrap samples with noise
  nm.stat <- replicate(R, statistic(add_noise(tseries[nmblk(n, l)])))
  mv.stat <- replicate(R, statistic(add_noise(tseries[mvblk(n, l)])))

  # Calculate p-values
  nmp.value <- (sum(nm.stat >= stat) + 0.5) / (R + 1)
  mvp.value <- (sum(mv.stat >= stat) + 0.5) / (R + 1)

  list(stat = stat, nmp.value = nmp.value, mvp.value = mvp.value)
}

# Calculate ks-statistic
ks_statistic <- function(x) ks.test(x, "pnorm")$statistic

# 1 test
do1rep <- function(n, ar, l) {

  # Create correlated data
  x <- arima.sim(list(ar = ar), rand.gen = rnorm, sd = sqrt(1-ar^2), n)

  # Apply block bootstrap
  bt <- blkboot(x, ks_statistic, R = 1000, l)

  list(stat = bt$stat, nmp.value = bt$nmp.value, mvp.value = bt$mvp.value)
}

# Naive example
test_naive_block <- function(ar, n) {

  # Generate correlated data
  x <- arima.sim(list(ar = ar), rand.gen = rnorm, sd = sqrt(1-ar^2), n)

  # Gather p-value
  p.value <- ks.test(x, "pnorm")$p.value

  list(p.value = p.value)
}

n <- 100
ar <- 0.5
l = 5
n_tests <- 100

# Generate results
pvals_block <- replicate(n_tests, do1rep(n, ar, l)$mvp.value)
pvals_naive_block <- replicate(n_tests, test_naive_block(ar = ar, n = n)$p.value)

# Create a master data_frame with all the results
histogram <- c(rep("Block", n_tests), rep("Naive", n_tests))
p_values <- c(pvals_block, pvals_naive_block)
results <- data_frame(p_values, histogram)

# Generate histograms plot
hist_block <- ggplot(results) +
             geom_histogram(aes(x = p_values, fill = histogram), boundary = 0, binwidth = 1/20) + 
             scale_x_continuous(name = "p-value", breaks=c(0, 1)) + 
             facet_grid(~histogram) +
             theme(legend.position = "none")

# Save the histogram figure
ggsave(filename = "hist_block.pdf", plot = hist_block, path = "../Manuscript", height = 2.1, width = 6)

proc.time() - ptm
```

```{r snpar}
test_snpar <- function(ar, n) {

  # Generate correlated data
  x <- arima.sim(list(ar = ar), rand.gen = rnorm, sd = sqrt(1-ar^2), n)

  # Gather p-value
  p.value <- snpar::KS.test(x, "pnorm")$p.value

  list(p.value = p.value)
}

n <- 100
ar <- 0.5
n_tests <- 1000

# Generate results
pvals_snpar <- replicate(n_tests, test_snpar(ar = ar, n = n)$p.value)
pvals_naive_block <- replicate(n_tests, test_naive_block(ar = ar, n = n)$p.value)

# Create a master data_frame with all the results
histogram <- c(rep("snpar", n_tests), rep("Naive", n_tests))
p_values <- c(pvals_snpar, pvals_naive_block)
results <- data_frame(p_values, histogram)

# Generate histograms plot
hist_snpar <- ggplot(results) +
             geom_histogram(aes(x = p_values, fill = histogram), boundary = 0, binwidth = 1/20) + 
             scale_x_continuous(name = "p-value", breaks=c(0, 1)) + 
             facet_grid(~histogram) + 
             theme(legend.position = "none")
```

@@

# Create the histograms displaying p-values from each test
hist1 <- suppressWarnings(test_rounded(1000, 100, 2))
hist2 <- suppressWarnings(test_rounded(1000, 100, 1))

# Create a master data_frame with all the results
histogram <- c(rep(".01", 1000), rep(".1", 1000))
p_values <- c(hist1, hist2)
results <- data_frame(p_values, histogram)

@@

```{r hist_general}
# Test the KSgeneral package as a solution to rounded data
test_general <- function(n_tests, n_samples, digits, B = 100) {

@@

  for (i in 1:n_tests) {

    # Generate random rounded data, perform 3 KS-tests, and store the p-values
    # 1. Using the rounded normal step function
    # 2. Using the naive continuous standard normal function
    data <- round(rnorm(n_samples), digits)
    results1[i] <- KSgeneral::disc_ks_test(data, sfun, exact = TRUE)$p.value
    results2[i] <- dgof::ks.test(data, sfun)$p.value
    results3[i] <- ks.test(data, sfun, simulate.p.value = TRUE, B = B)$p.value
    results4[i] <- ks.test(data, "pnorm")$p.value

  }

@@

n_tests = 100
n = 50

# Create the histograms displaying p-values from each test
test1 <- suppressWarnings(test_general(n_tests, n, 2))
test2 <- suppressWarnings(test_general(n_tests, n, 1))
results <- rbind(test1, test2)

@@

```{r babu.ts}
ks.np.babu.ts <- function(x, l, B = 1000) {
  ## for normal distribution only
  n <- length(x)
  Fn <- ecdf(x)
  Gn <- ecdf.leftcont(x)
  mu <- mean(x); ss <- sd(x)
  Ftheta <- pnorm(x, mu, ss)
  Bn1 <- Fn(x) - Ftheta
  Bn2 <- Gn(x) - Ftheta
  d1 <- abs(Bn1)
  d2 <- abs(Bn2)
  stat <- max(d1, d2)
  stat.b <- double(B)
  for (i in 1:B) {
    x.b <- x[nmblk(n, 5)]
    Fn.b <- ecdf(x.b)
    Gn.b <- ecdf.leftcont(x.b)
    mu.b <- mean(x.b); ss.b <- sd(x.b)
    ## evaluate at x (not x.b)
    Ftheta.b <- pnorm(x, mu.b, ss.b)
    d1.b <- abs(Fn.b(x) - Ftheta.b - Bn1)
    d2.b <- abs(Gn.b(x) - Ftheta.b - Bn2)
    stat.b[i] <- max(d1.b, d2.b)
  }
  p.value <-  (sum(stat.b >= stat) + 0.5) / (B + 1)
  list(p.value = p.value,
       statistic = stat, stat.b = stat.b)
}
```

```{r hist_babu}
test_babu <- function(n_tests, n, ar, l) {

  results1 <- double(n_tests)
  results2 <- double(n_tests)
  results3 <- double(n_tests)

  for (i in 1:n_tests) {

    cat(i, " ")

    x <- arima.sim(list(ar = ar), rand.gen = rnorm, sd = sqrt(1-ar^2), n = n)

    results1[i] <- ks.np.babu.ts(x, l, B = 1000)$p.value
    results2[i] <- blkboot(x, ks_statistic, R = 1000, l)$nmp.value
    results3[i] <- ks.test(x, "pnorm")$p.value

  }

  # Create a master data_frame for p-values
  test <- c(rep("Babu", n_tests), rep("Block", n_tests), rep("Naive", n_tests))
  p_values <- c(results1, results2, results3)
  results <- data_frame(p_values, test)

n_tests = 100
n = 100
l = 10
ar = 0.5

# Create the histograms displaying p-values from each test
results <- suppressWarnings(test_babu(n_tests = n_tests, n = n, ar = ar, l = l))

# Time per test
(proc.time() - ptm)["elapsed"] / n_tests

# Total time
proc.time() - ptm

# Generate histograms plot
hist_babu <- ggplot(results) + 
                geom_histogram(aes(x = p_values, fill = test), boundary = 0, binwidth = 1/20) + 
                scale_x_continuous(name = "p-value", breaks=c(0, 1)) + 
                facet_grid(~test) +
                theme(legend.position = "none")
hist_babu
```

```{r}
#hist_babu1000
#hist_babu_1000
```

>>>>

```{r}
# Load packages
library(xtable)
library(dgof)
library(tidyverse)
library(KSgeneral)
library(copula)
```

@@

  # Parametric bootstrap to approximate the null distribution
  n <- length(x)
  stat.b <- double(B)
  for (i in 1:B) {
    x.b <- rnorm(n, mu, sigma)
    mu.b <- mean(x.b)
    sigma.b <- sd(x.b)
    stat.b[i] <- ks.test(x.b, "pnorm", mu.b, sigma.b)$statistic
  }
  p.value <- (sum(stat.b >= stat) + 0.5) / (B + 1)

  counter <<- counter + 1
  cat(counter, "")
  list(p.value = p.value)
}

test_babu <- function(x, B = 1000) {
  p.value <- ks.np.babu(x, B = 1000)$p.value
  counter <<- counter + 1
  cat(counter, "")
  list(p.value = p.value)
}

@@

# Generate results
n_tests = 100
n = 100
counter <- 0
pvals_parametric <- replicate(n_tests, test_parametric(rnorm(n), B = 1000)$p.value)
counter <- 0
pvals_babu <- replicate(n_tests, test_babu(rnorm(n), B = 1000)$p.value)
pvals_naive <- replicate(n_tests, test_naive(rnorm(n))$p.value)

# Create a master data_frame with all the results
histogram <- c(rep("Parametric", n_tests), rep("Babu", n_tests), rep("Naive", n_tests))
p_values <- c(pvals_parametric, pvals_babu, pvals_naive)
results <- data_frame(p_values, histogram)

# Generate histograms plot
hist_fitted <- ggplot(results) + 
                   geom_histogram(aes(x = p_values, fill = histogram), boundary = 0, binwidth = 1/20) + 
                   scale_x_continuous(name = "p-value", breaks=c(0, 1)) + 
                   theme(legend.position = "none") +
                   facet_wrap(~histogram, scales = "free_y")

# Save the histogram figure
ggsave(filename = "hist_fitted.pdf", plot = hist_fitted, path = "../Manuscript", height = 2.1, width = 6)
```

@@

# Create the histograms displaying p-values from each test
n_tests <- 1000
n <- 100
hist1 <- suppressWarnings(test_rounded(n_tests, n, 2))
hist2 <- suppressWarnings(test_rounded(n_tests, n, 1))

# Create a master data_frame with all the results
histogram <- c(rep(".01", n_tests), rep(".1", n_tests))
p_values <- c(hist1, hist2)
results <- data_frame(p_values, histogram)

@@

```{r hist_general}
# Test the KSgeneral package as a solution to rounded data
test_general <- function(n_tests, n_samples, digits, B) {

@@

  for (i in 1:n_tests) {
    cat(i, "")
    # Generate random rounded data, perform 3 KS-tests, and store the p-values
    # 1. Using the rounded normal step function
    # 2. Using the naive continuous standard normal function
    data <- round(rnorm(n_samples), digits)
    results1[i] <- KSgeneral::disc_ks_test(data, sfun, exact = TRUE)$p.value
    results2[i] <- dgof::ks.test(data, sfun)$p.value
    results3[i] <- ks.test(data, sfun, simulate.p.value = TRUE, B = B)$p.value
    results4[i] <- ks.test(data, "pnorm")$p.value
  }

@@

n_tests = 100
n = 100
B = 1000

# Create the histograms displaying p-values from each test
test1 <- suppressWarnings(test_general(n_tests, n, digits = 2, B = B))
test2 <- suppressWarnings(test_general(n_tests, n, digits = 1, B = B))
results <- rbind(test1, test2)

@@

```{r}
test_copula <- function(x, dist, B = 100) {
    ## example assuming normal distribution; normal copula
    ## get observed stat
    n <- length(x)
    mu <- mean(x)
    sigma <- sd(x)
    stat  <- ks.test(x, dist, mean = mu, sd = sigma)$statistic
    ## get lag-1 sample auto-spearman rho
    rho  <-  cor(x[-1], x[-n], method = "spearman")
    r <- iRho(normalCopula(), rho)
    np <- normalCopula(r)
    ## parametric bootstrap to get an empirical distribution of stat
    stat.b <- double(n)
    ## set up containers
    for (i in 1:B) {
        ## get one bootstrap sample
        ## 1. the marginal distribution is the fitted dist
        ## 2. there is temporal dependence
        u <- runif(n)
        for (j in 2:n) {
            ## could test other copulas later
            u[j]  <- cCopula(c(u[j-1], u[j]), np, indices = 2, inverse = TRUE)
        }
        ## get bootstrap sample
        x.b <- qnorm(u, mean = mu, sd = sigma)
        ## fit it
        mu.b <- mean(x.b)
        sigma.b <- sd(x.b)
        stat.b[i] <- ks.test(x.b, dist, mean = mu.b, sd = sigma.b)$statistic
    }

    p.value <- (sum(stat.b >= stat) + 0.5) / (B + 1)

    counter <<- counter + 1
    cat(counter)

    return(p.value)
    # return empirical p-value    
}

# Naive KS-test
test_naive_copula <- function(x) {
  p.value <- ks.test(x, "pnorm")$p.value
  list(p.value = p.value)
}

# Generate results
n <- 100
ar = 0.5
counter <- 0
nrep <- 1000
pvals_copula <- replicate(nrep, test_copula(arima.sim(list(ar = ar), rand.gen=rnorm, sd = sqrt(1 - ar^2), n = n), "pnorm"))
pvals_naive <- replicate(nrep, test_naive_copula(arima.sim(list(ar = ar), rand.gen=rnorm, sd = sqrt(1 - ar^2), n = n))$p.value)

# Create a master data_frame with all the results
histogram <- c(rep("Copula", nrep), rep("Naive", nrep))
p_values <- c(pvals_copula, pvals_naive)
results <- data_frame(p_values, histogram)

# Generate histograms plot
hist_copula <- ggplot(results) + 
                   geom_histogram(aes(x = p_values, fill = histogram), boundary = 0, binwidth = 1/20) + 
                   scale_x_continuous(name = "p-value", breaks=c(0, 1)) + 
                   theme(legend.position = "none") +
                   facet_wrap(~histogram, scales = "free_y")
hist_copula

# Save the histogram figure
ggsave(filename = "hist_copula.pdf", plot = hist_copula, path = "../Manuscript", height = 2.1, width = 6)
```
